Spin canonical rings of log stacky curves

### 2015-07-11 surfaces-write-up/surface-write-up.tex
... I didn't think I made any changes

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\bbf{{\mathbb F}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}
\newcommand\bk{{\Bbbk}}

@@

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d,c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_r \in \Sigma,$ with $e_i = (\delta^i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_r$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$. Furthermore, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r} w_i e_i,$ with the first coordinate of $\lambda$ less than $\sum_{i=1}^{r}\delta_i$ and $w_1,\ldots, w_r \in \bz$.
\end{prop}
\begin{proof}
By assumption, $\sigma \in \Sigma$ as $\sigma = \sum_{i=1}^{r}a_i e_i$ with $a_i \in \br$. Let $\{a\} := a - \lfloor a \rfloor$ denote the fractional part of $a$. Let $\lambda = \sum_{i=1}^{r}\{a_i\}$. Hence, we can write $\sigma = \lambda + \sum_{i=1}^{r}\lfloor a_i \rfloor e_i.$ Hence, $\sigma$ lies in the $\bz_{\geq 0}$ span of $\lambda, e_1,\ldots, e_r$, which all have first coordinate less than $\sum_{i=1}^{r}\delta_i$. Therefore, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{r}\delta_i$.
\end{proof}

\begin{lem}
\label{lem:composite-map}
Let $D = \sum_{i=1}^{n}\alpha_i D_i$ where $D_i = V(f_i)$. Suppose we have a surjection $\phi: \bk[x_1,x_2,\ldots, x_r] \rightarrow R_D, x_i \mapsto p_i(f_1, \ldots, f_n),$ where $p_i$ is a monomial in $f_1,\ldots, f_n$. Then, define
\begin{align*}
	\Sigma = \langle u^d y_1^{c_1} \cdots y_n^{c_n} : c_i \geq -d \alpha_i, \sum_{i=1}^{n}a_ic_i = \sum_{i=1}^{} b_i c_i. \rangle 
\end{align*}
In this case, we have a factorization of $\phi$ given by maps $\chi, \psi$ defined by
\todo{Make this have mapsto on bottom}
$$\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}y_1^{c_{i1}} \cdots y_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}
\end{tikzcd}$$
Assuming $\chi$ is surjective, the minimal degree of generation of $\ker \phi$ is at most the maximum of the minimal degree of generation of $\ker \chi$ and the minimal degree of generation of $\ker \psi$.
\end{lem}
\begin{proof}
Note that we have an exact sequence
$$\begin{tikzcd}
0 \ar r & \ker \chi \ar r & \ker \phi \ar r & \ker \psi \ar r & 0
\end{tikzcd}.$$
To see this, note that the left map is an inclusion because $\phi = \chi \circ \psi$. Next, the right map, given by $\chi|_{\ker \phi}:\ker \phi \rightarrow \ker \psi$ is a surjection, again because $\phi = \chi \circ \psi$. The composition of the two maps is $0$ by definition of the kernel of $\chi$. Finally, if $a \in \ker \phi$ and $a$ lies in the kernel of the right map, then by definition $\chi(a) = 0$.
\end{proof}

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has extremal rays $e_1,\ldots, e_r$ in degrees $d_1, \ldots, d_r$ then $\ker \chi$ has a system of generators in degrees less than $\sum_{i=1}^{r}d_r-1$.
\end{lem}
\begin{proof}
Since $e_1, \ldots, e_r$ are extremal rays, by Proposition ~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r}e_i$. Now, letting $\lambda_0 := 0,\lambda_1, \ldots, \lambda_m$ be all elements of $\Sigma$ so that we can write $\lambda_i = \sum_{i=1}^{r}a_i e_i$ with $0 \leq a_i < 1$ for all $i$. Then, for any $1 \leq i \leq j \leq m,$ we can write $\lambda_i + \lambda_j$ in the above canonical form, yielding a relation in degree at most $\deg \lambda_i + \deg \lambda_j \leq 2 \cdot \left( \sum_{i=1}^{r}d_r -1 \right).$ Furthermore, these relations generate all relations, as one can apply a sequence of these relations to put any $\sigma \in \Sigma$ into canonical form.
\end{proof}

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = d_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. We shall further assume that among there exist functions $f_0,\ldots, f_{m+1}$ form a basis for degree one rational functions on $\bp^m$.

\todo{possibly change this to only have two equations}
\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ so that 
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot d_i \cdot f_i = 0 & \text{ and } &c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i} $ such that
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor & \text{ for } & i > m
\end{align}
form a basis for $H^0(\bp^m, dD)$ over $k$.
\end{prop}
\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}
\todo{the following bound is the generation bound beginning from evan's paper}

@@

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{Robin: do this}
\end{rem}

@@

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ generators $f_{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $\vec{v} \in \bz_{\geq 0}^{m + 1}$
such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
relations of the form
\[
	x^{\vec{u}} x^{\vec{v}} = x^{\vec{w}} x^{\vec{z}}
\]

@@

\section{Canonical Rings on \texorpdfstring{$F_k$}{}}
\begin{defn}
Suppose $D = \sum_{i=1}^m \alpha_i C_i \in \mathbb{Q} \otimes \di(F_k)$.  Then define 
\begin{equation}\label{eqn:define-T=(D)}
	T_=(D) = \left\{i \in \{1, \ldots, m\}: d_i \sum_{k=1}^m d_k' \alpha_k = d_i' \sum_{k=1}^m d_k \alpha_k \right\},
\end{equation}
\begin{equation}\label{eqn:define-T+(D)}
	T_+(D) = \left\{ i \in \{1, \ldots, m\}:  \vphantom{\sum_{k=1}^m} d_i \sum_{k=1}^m \alpha_k d_k' > d_i' \sum_{k=1}^m \alpha_k d_k \right\},
\end{equation}
and
\begin{equation}\label{eqn:define-T-(D)}
	T_-(D) = \left\{ j \in \{1, \ldots, m\}: d_j \sum_{k=1}^m \alpha_k d_k' < d_j' \sum_{k=1}^m \alpha_k d_k s \right\}.
\end{equation}
\end{defn}

\begin{lem}
If $D=\sum_{i=1}^m \alpha_i C_i$ for some curve $C_i\in F_k$, such that $C_1, \ldots, C_4$ are independent curves with bi-degrees (1,0), (1,0), (0,1), and (0,1) respectively.  Then each $C_i$ corresponds to a polynomial $f_i$ where $f_1$ and $f_2$ are independent linear polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent linear polynomials in $y_0$ and $y_1$.  Then $D$ is generated in degree at most
\[
	\sum_{i\in T_=(D)} \gcd(d_i, d_i')\ell_i + \sum_{i\in T_+(D) \atop j\in T_-(D)} (d_i d_j'- d_j d_i')\ell_{i,j}
\]
 with relations in degree at most \todo{add}.
\end{lem}
\begin{proof}
Let $(d_i, d_i')$ be the bi-degrees of $f_i$.

Notice that if $\gamma \in (R_D)_d$, then 
\[
	g = \sum_{i=1}^m {f_i}^{a_i}
\]
where each $a_i \ge - \alpha_i d_i$, $\sum_{i=1}^m a_i d = 0$, and $\sum_{i=1}^m a_i d_i' = 0$.  Then we can view $g$ inside of the lattice
$S\subseteq \mathbb{R}^{n+1}$ given by ordered elements of the form $(d, a_1, \ldots, a_n)$ satisfying $a_i \ge - d \alpha_i$ for all $i$, $\sum_{i=1}^m a_i d_i = 0$, and $\sum_{i=1}^m a_i d_i' = 0$. 
From here, determining a spanning set for $(R_D)_d$ reduces to finding the extremal rays of $S$ \todo{cite previous lemma}.

To do this, we extend the method of O'Dorney \todo{cite}.  We first view the cone $T\subseteq \mathbb{R}^{n+1}$ given by elements of the form $(d, a_1, \ldots, a_n)$ satisfying $a_i \ge -\alpha_i d$ for all $i$ and $\sum_{i=1}^n a_i (d_i + d_i') =0$.  By O'Dorney, this has extremal rays given by 
\[
	e_i = (1, -\alpha_1, \ldots, -\alpha_{i-1}, \frac{\sum_{j \ne i} \alpha_j(d_j+ d_j')}{d_i + d_i'}, -\alpha_{i+1}, \ldots, - \alpha_m).
\]
Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_m$.  We can then intersect the hyperplane $H$ given $\sum_{i=1}^m  d_i X_i = 0$.  The extremal rays of $S$ are given by the set of $e_i$ contained in $H$ together with the points of intersection $e_{i,j}$ of $H$ and the line segment connecting $e_i, e_j \not \in H$.

@@

For $i \in T_=(D)$
define
\[
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha \ell_i \gcd(d_i, d_i')}) (f_i^{\ell_i \frac{\gcd(d_i, d_i')}{d_i + d_i'}\sum_{k \ne i} \alpha_k (d_k + d_k')}) \in H^0(\ell_i \gcd(d_i, d_i') D).
\]
For $i \in T_+(D)$ and $j \in T_-(D)$, define
\[
	\epsilon_{i, j} = (\prod_{k = 1}^m f_k^{-\alpha \ell_{i,j} (d_i d_j' - d_j d_i')}) {f_i}^{s_1} {f_j}^{s_2} \in H^0(\ell_{i,j}(d_i d_j' - d_j d_i')D)
\]
where
\[
	s_1 = d_j' \sum_{k \ne i,j} \alpha_k d_k - d_j \sum_{k\ne i, j} \alpha_k d_k'
\]
and
\[
	s_2 = d_i' \sum_{k \ne i,j} \alpha_k d_k - d_i \sum_{k \ne i, j} \alpha_k d_k'.
\]
Let $E$ be the set of $\epsilon_i$ and $\epsilon_{i,j}$ given above, satisfying the respective conditions.
$E$ satisfies the conditions of proposition \todo{cite earlier proposition}, so by applying that proposition $R_D$ is generated in degree at most
\[
	\sum_{i\in T_=(D)} \gcd(d_i, d_i')\ell_i + \sum_{i\in T_+(D) \atop j\in T_-(D)} (d_i d_j'- d_j d_i')\ell_{i,j}.
\]

>>>>

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\bbf{{\mathbb F}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}

@@

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d,c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_r \in \Sigma,$ with $e_i = (\delta^i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_n$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$.
\end{prop}
\begin{proof}
By assumption, $\sigma \in \Sigma$ as $\sigma = \sum_{i=1}^{n}a_i e_i$ with $a_i \in \br$. Let $\{a\} := a - \lfloor a \rfloor$ denote the fractional part of $a$. Let $\lambda = \sum_{i=1}^{n}\{a_i\}$. Hence, we can write $\sigma = \lambda + \sum_{i=1}^{n}\lfloor a_i \rfloor e_i.$ Hence, $\sigma$ lies in the $\bz_{\geq 0}$ span of $\lambda, e_1,\ldots, e_n$, which all have first coordinate less than $\sum_{i=1}^{n}\delta_i$. Therefore, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$.
\end{proof}

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = a_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. We shall further assume that among there exist functions $f_0,\ldots, f_{m+1}$ form a basis for degree one rational functions on $\bp^m$.

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ so that 
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot a_i \cdot f_i = 0 & \text{ and } &c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i} $ such that
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor & \text{ for } & i > m
\end{align}
form a basis for $H^0(\bp^m, dD)$ over $k$.
\end{prop}
\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{do this}
\end{rem}

@@

@@

\noindent
has a minimal presentation consisting of the $\sum_{i = 0}^{r}
{{m + i} \choose {i}}$ generators $f_{\vec{v}} := \frac{u^{a_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $\vec{v} \in \bz_{\geq 0}^{m + 1}$
such that $\sum_{j = 0}^{m} v_j = c_i$ and $z$ \todo{how many?}
relations of the form
\[
	x^{\vec{u}} x^{\vec{v}} = x^{\vec{w}} x^{\vec{z}}
\]

@@

\section{Canonical Rings on \texorpdfstring{$F_k$}{}}
Let $D=\sum_{i=1}^m \alpha_i C_i$ for some curve $C_i\in F_k$, such that $C_1, \ldots, C_4$ are independent curves with bi-degrees (1,0), (1,0), (0,1), and (0,1) respectively.  Let $C_i = V(f_i)$ where $f_i \in \mathscr{O}(a_i, b_i)$.  Then each $C_i$ corresponds to a polynomial $f_i$ where $f_1$ and $f_2$ are independent linear polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent linear polynomials in $y_0$ and $y_1$.
\begin{defn}
Suppose $D = \sum_{i=1}^m \alpha_i C_i \in \mathbb{Q} \otimes \di(F_k)$.  Then define 
\begin{equation}\label{eqn:define-T=(D)}
	T_=(D) = \left\{i \in \{1, \ldots, m\}: a_i \sum_{k=1}^m b_k \alpha_k = b_i \sum_{k=1}^m a_k \alpha_k \right\},
\end{equation}
\begin{equation}\label{eqn:define-T+(D)}
	T_+(D) = \left\{ i \in \{1, \ldots, m\}:  \vphantom{\sum_{k=1}^m} a_i \sum_{k=1}^m \alpha_k b_k > b_i \sum_{k=1}^m \alpha_k a_k \right\},
\end{equation}
and
\begin{equation}\label{eqn:define-T-(D)}
	T_-(D) = \left\{ j \in \{1, \ldots, m\}: a_j \sum_{k=1}^m \alpha_k b_k < b_j \sum_{k=1}^m \alpha_k a_k s \right\}.
\end{equation}
\end{defn}

\begin{lem}
$R_D$ is generated in degree at most
\[
	\sum_{i\in T_=(D)} \gcd(a_i, b_i)\ell_i + \sum_{i\in T_+(D) \atop j\in T_-(D)} (a_i b_j- a_j b_i)\ell_{i,j}.
\]
\end{lem}
\begin{proof}
Let $(a_i, b_i)$ be the bi-degrees of $f_i$.

Notice that if $\gamma \in (R_D)_d$, then 
\[
	g = \sum_{i=1}^m {f_i}^{c_i}
\]
where each $c_i \ge - \alpha_i c_i$, $\sum_{i=1}^m c_i a_i = 0$, and $\sum_{i=1}^m c_i b_i = 0$.  Then we can view $g$ inside of the lattice
$S\subseteq \mathbb{R}^{n+1}$ given by ordered elements of the form $(d, c_1, \ldots, c_n)$ satisfying $c_i \ge - d \alpha_i$ for all $i$, $\sum_{i=1}^m c_i a_i = 0$, and $\sum_{i=1}^m c_i b_i = 0$. 
From here, determining a spanning set for $(R_D)_d$ reduces to finding the extremal rays of $S$ \todo{cite previous lemma}.

To do this, we extend the method of O'Dorney \todo{cite}.  We first view the cone $T\subseteq \mathbb{R}^{n+1}$ given by elements of the form $(d, c_1, \ldots, c_n)$ satisfying $c_i \ge -\alpha_i d$ for all $i$ and $\sum_{i=1}^n c_i (a_i + b_i) =0$.  By O'Dorney, this has extremal rays given by 
\[
	e_i = (1, -\alpha_1, \ldots, -\alpha_{i-1}, \frac{\sum_{j \ne i} \alpha_j(a_j+ b_j)}{a_i + b_i}, -\alpha_{i+1}, \ldots, - \alpha_m).
\]
Let $\Delta^{n-1}$ be the simplex with vertices $e_1, \ldots e_m$.  We can then intersect the hyperplane $H$ given $\sum_{i=1}^m  a_i X_i = 0$.  The extremal rays of $S$ are given by the set of $e_i$ contained in $H$ together with the points of intersection $e_{i,j}$ of $H$ and the line segment connecting $e_i, e_j \not \in H$.

@@

For $i \in T_=(D)$
define
\[
	\epsilon_i := (\prod_{k \ne i} f_k^{-\alpha \ell_i \gcd(a_i, b_i)}) (f_i^{\ell_i \frac{\gcd(a_i, b_i)}{a_i + b_i}\sum_{k \ne i} \alpha_k (a_k + b_k)}) \in H^0(\ell_i \gcd(a_i, b_i) D).
\]
For $i \in T_+(D)$ and $j \in T_-(D)$, define
\[
	\epsilon_{i, j} = (\prod_{k = 1}^m f_k^{-\alpha \ell_{i,j} (a_i b_j - a_j b_i)}) {f_i}^{s_1} {f_j}^{s_2} \in H^0(\ell_{i,j}(a_i b_j - a_j b_i)D)
\]
where
\[
	s_1 = b_j \sum_{k \ne i,j} \alpha_k a_k - a_j \sum_{k\ne i, j} \alpha_k b_k
\]
and
\[
	s_2 = b_i \sum_{k \ne i,j} \alpha_k a_k - a_i \sum_{k \ne i, j} \alpha_k b_k.
\]
Let $E$ be the set of $\epsilon_i$ and $\epsilon_{i,j}$ given above, satisfying the respective conditions.
$E$ satisfies the conditions of proposition \todo{cite earlier proposition}, so by applying that proposition $R_D$ is generated in degree at most
\[
	\sum_{i\in T_=(D)} \gcd(a_i, b_i)\ell_i + \sum_{i\in T_+(D) \atop j\in T_-(D)} (a_i b_j- a_j b_i)\ell_{i,j}.
\]
\end{proof}

\todo{factor $\phi$ through $\psi$ and $\chi$.}
\begin{lem}
$\ker(\psi)$ is generated in degree at most \todo{insert number}.
\end{lem}
\begin{proof}
We first claim that $\psi$ has a relation of the form
\[
	(y_j - \beta_j)\prod_{i=1}^m {y_i}^{c_{i}}
\]
whenever $d\in \mathbb{N}$, $c_i \ge -\alpha_i d$, and $\beta_j \in k[y_1, \ldots, y_4]$ satisfies $\psi(\beta_j)\in \mathscr{O}(a_j, b_j)$.
To find such a $\beta$, we simply write $\epsilon_j$ as a polynomial in $\epsilon_1, \ldots \epsilon _4$.
